Scaffolder - Software for Microbial Genome Scaffolding.

### 2011-02-01 text/src/article.tex
Revise third section of the results

Scaffolder provides a ruby API that allows the scaffold and designated
sequences to be manipulated through a ruby programme. This allows scaffolder to
be integrated into exiting genomics works flows such as using Rake.  Scaffolder
all provides a command line binary for the manipulation of the scaffold. This
binary behaves as a standard Unix command line interface returning appropriate
exit statuses and providing manual pages for each of the commands. These two
methods are outlined in the following three sections. \pb

@@

Scaffolder relies on both the ruby programming language and the RubyGems
package management system. These are commonly installed on recent versions of
Mac OSX and Linux distributions. The installation of these can be tested on the
command line by typing the following commands. \pb

@@

If both these commands return without error this indicates the necessary
requirements to install scaffolder. The scaffolder API and command line tools
can then be installed with the follow command.

@@

Finally to ensure that the command line tools are accessible these the
RubyGems binary directory should be in the command line path. This can be done
by adding the following to the ~/.bashrc file or other corresponding shell
configure file. Assuming that ~/.gem is the install path for RubyGems.

@@

\subsubsection*{Command line interface} %{{{3

Scaffolder provides a command line interface to validate and build a genome
sequence from a scaffold. Assuming scaffolder has been installed and the
binary is available in the command line path the following command can be used
to invoke scaffolder. This will list the available commands for running
scaffolder. \pb

\begin{verbatim}
$ scaffolder
\end{verbatim}

Help on each command can be found by type `help' followed by the name of the
command. This will return the Unix manual page for the command of interest.
For example the following command will return the manual page for sequence
listing the available options for running the sequence command. \pb

\begin{verbatim}
$ scaffolder help sequence
\end{verbatim}

To the generate the fasta sequence from a scaffold the sequence command should
be used. This should be passed the name of the file with containing the YAML
formatted scaffold and the corresponding fasta file of sequences. This will
then return the complete drafted sequence to the standard out. \pb

\begin{verbatim}
$ scaffolder sequence scaffold.yml sequences.fna
\end{verbatim}

Scaffolder also provides a function to test the scaffold for overlapping
insert sequence co-ordinates. As described in the section above overlapping
inserts will result in unexpected behaviour as the coordinates for the first
insert will change the expected insert location for the second insert. The
validate command will check the scaffold for any overlapping inserts and
return them as a YAML formatted list describing each one. \pb

\begin{verbatim}
$ scaffolder validate scaffold.yml sequences.fna
\end{verbatim}

@@

The scaffolder tool has a defined application programming interface (API) to
allow ruby programs to manipulate the scaffold. Detailing the complete API
would be overwhelming to this document. The scaffolder API is fully documented
online using at the scaffolder website. The following code snippet gives and
outline of how the API may be used to create a script to replicate the
function of the `scaffolder sequence' command described above. \pb

@@

This program illustrates that the scaffold object is initialised with two arguments: the YAML loaded scaffold file and the location of the fasta formatted sequence file. The scaffold is then enumerable as the next line illustrates the trivial code required to generate the scaffold. Further details of the attributes for each entry in the scaffold can be found in the documentation for the Scaffolder::Region class. \pb

>>>>

Scaffolder provides a standardised set of Ruby classes and methods (the
application programming interface or API) allowing the scaffold to be
manipulated programmatically. This allows scaffolder to be integrated into
existing genomics workflows or used with Ruby build tools such as using Rake.
\pb

Scaffolder also provides a command line interface (CLI) through a scaffolder
binary. This the scaffold sequence to be produced and validated without
requiring any Ruby knowledge. The scaffolder CLI behaves as a standard Unix
tool - returning appropriate exit statuses and providing appropriate manual
pages. The use of both these scaffolder interfaces is outlined in the following
sections. \pb

@@

Scaffolder relies on both the ruby programming language and the RubyGems
package management system to work. These are either already present easily
installed on recent distributions of Mac OSX or Linux. The presences of these
tools can be tested by typing the following on the command line. \pb

@@

If both these commands return without error this indicates the necessary
system requirements to install and run scaffolder. The scaffolder API and CLI
can then be installed through RubyGems with the following command.

@@

To ensure that the scaffolder CLI is accessible the RubyGems `bin' directory
should be in the command line path. Assuming that RubyGems are installed in
~/.gem this can be done by adding the following to the ~/.bashrc file. This
should be adjusted for other shells such as zsh etc..

@@

\subsubsection*{Command Line Interface} %{{{3

The Scaffolder CLI provides functions to validate the scaffold file and build the draft scaffold sequence. Assuming scaffolder has been installed and the
binary is available in the command line path the following command should run without error and list the available functions. \pb

\begin{verbatim}
  scaffolder
\end{verbatim}

Help on a specific command can be found by typing `help' followed by the name
of the command. This will return the corresponding Unix manual page. The
following example will return the manual page for the sequence command. \pb

\begin{verbatim}
  scaffolder help sequence
\end{verbatim}

The ``sequence'' command builds the fasta sequence from the scaffold and fasta
file. Running this command requires the locations of both the scaffold file and
the fasta file of sequences. This will then return the complete scaffold
sequence. \pb

\begin{verbatim}
  scaffolder sequence scaffold.yml sequences.fna
\end{verbatim}

Scaffolder provides a function to test the scaffold for overlapping insert
coordinates. Overlapping inserts will result in unexpected behaviour as the
coordinates for the first insert may change the insert location for subsequent
inserts. The ``validate'' command will check the scaffold for any such inserts
and return them as a YAML list. \pb

\begin{verbatim}
  scaffolder validate scaffold.yml sequences.fna
\end{verbatim}

@@

Scaffolder provides an API allowing Ruby programs to manipulate the scaffold.
This API is detailed in full on the scaffolder website. The following code
snippet outlines how the API may be used to create a script to replicate the
function of the `scaffolder sequence' command described above. \pb

@@

This code snippet illustrates that the scaffold object is initialised with two
arguments: the YAML loaded scaffold and the location of the fasta sequence
file. The scaffold is an enumerable object as the next line illustrates the
trivial code required to generate the scaffold sequence. Further details of
the classes and method related to each scaffold entry can be found in the
documentation for the Scaffolder::Region class. \pb
